---
title: "Calc_USDA_NASS_Crops_HUC12_Raster_Covariates"
author: "Kyle P Messier"
date: "5/23/23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required Libraries and Source Local Functions

```{r libraries and sourcing,echo=FALSE}
library(sf)
library(terra)
library(exactextractr)
library(lubridate)
library(tidyverse)
library(data.table)
library(nhdplusTools)
library(stringr)
```

## Prepare Pesticide Data - sf format

```{r pesticide and NHD WBD data, echo=FALSE}
# Read in main pesticide data here
data.AZO <- sf::st_read("/Volumes/messierkp/Projects/CensoredGP/WaterData-Pesticides/nwis_triazines_alachlors/data_AZO_watershed_huc_join.shp")

# For efficient extraction, we just need the geometry
AZO.geometry <- sf::st_geometry(data.AZO)

# US bounding box
US.bb <- terra::ext(c(-124.7844079, -66.9513812, 24.7433195, 49.3457868))

# NHD WBD Layer names
WBD.layers <- st_layers("/Volumes/SHAG/WBD-National/WBD_National_GDB.gdb")

WBD <- st_read("/Volumes/SHAG/WBD-National/WBD_National_GDB.gdb", layer = "WBDHU12")
```


## Read in the OpenLandMaps (OLM) soil raster data

Here, we have an if-else statement. If the cropped (to US bounding box) have been
created, then we read those. Else, we read in the raw data and then crop. The process
of cropping takes a bit of time - about an hour - so we want to only do that once
if we can help it. 

We use the terra package to read and process rasters, which is the most efficient
method currently available in R for dealing with rasters.

```{r NASS CropScape raster data,echo=FALSE}
# Cropped Raster directory


# Directories of individual raw OLM data
NASS2008 <- "/Volumes/SHAG/USDA_NASS/2008_30m_cdls.tif"
NASS2009 <- "/Volumes/SHAG/USDA_NASS/2009_30m_cdls.tif"
NASS2010 <- "/Volumes/SHAG/USDA_NASS/2010_30m_cdls.tif"
NASS2011 <- "/Volumes/SHAG/USDA_NASS/2011_30m_cdls.tif"
NASS2012 <- "/Volumes/SHAG/USDA_NASS/2012_30m_cdls.tif"
NASS2013 <- "/Volumes/SHAG/USDA_NASS/2013_30m_cdls.tif"
NASS2014 <- "/Volumes/SHAG/USDA_NASS/2014_30m_cdls.tif"
NASS2015 <- "/Volumes/SHAG/USDA_NASS/2015_30m_cdls.tif"
NASS2016 <- "/Volumes/SHAG/USDA_NASS/2016_30m_cdls.tif"
NASS2017 <- "/Volumes/SHAG/USDA_NASS/2017_30m_cdls.tif"
NASS2018 <- "/Volumes/SHAG/USDA_NASS/2018_30m_cdls.tif"
NASS2019 <- "/Volumes/SHAG/USDA_NASS/2019_30m_cdls.tif"


NASS_stack <- rast(c(NASS2008, NASS2009, NASS2010, NASS2011, NASS2012, NASS2013, NASS2014, NASS2015, NASS2016, NASS2017, NASS2018, NASS2019))


crop.classes <- levels(NASS_stack[[12]]$Class_Names) %>% as.data.frame()
# Rename a few that are duplicate names
crops.classes$Class_Names[65] <- "Non-Crop-Shrubland"
crops.classes$Class_Names[66] <- "Non-Crop-Barren"

for (i in 1:12) {
  levels(NASS_stack[[i]]) <- list(data.frame(
    ID = crop.classes$value,
    crop_name = crop.classes$Class_Names
  ))
}

year.range <- c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019)
names(NASS_stack) <- paste0("crops", year.range)

crop.classes$value <- as.character(crop.classes$value)
```

Here, we use the terra extract function to extract the exact raster value
at each of our pesticide locations. First, we have to convert the pesticide 
data to a SpatVect (terra package), a requirement of the extract function. 

```{r Calculate exact grid points ,echo=TRUE}
# HUC12

huc12.unique <- unique(data.AZO$huc12)


HUC.rast.class <- data.table("huc12" = data.AZO$huc12)


# need to drop and resolve names/layers
class.possible.names <- paste0("frac_", levels(NASS_stack)[[1]]$ID)

convert.name <- paste0("frac_", crops.classes$Class_Names)


HUC.rast.class[, paste0("huc12", ".", class.possible.names) := 0]



for (i in 1:length(huc12.unique)) {
  # for (i in 1:2){
  print(i)

  # get index - for rows of output - where data match the given HUC12
  idx.row <- which(data.AZO$huc12 == huc12.unique[i])

  # get the given HUC12 geometry from the WBD data
  huc12.polygon <- dplyr::filter(WBD, huc12 == huc12.unique[i])

  # calculate the fraction of each raster class in the HUC and by year
  # Note "frac" does not do stack applies, so we use an index to summarize
  extract.raster.classes <- exact_extract(NASS_stack, huc12.polygon, fun = "frac", stack_apply = TRUE) %>% t()

  # Create a dataframe with index to summarize, keep track of the crop value for
  # placement in the output
  str.idx <- rep(NA_real_, length(extract.raster.classes))
  crop.val <- rep(NA_real_, length(extract.raster.classes))
  counter <- 1
  for (j in 1:length(crop.classes$value)) {
    idx <- stringr::str_detect(rownames(extract.raster.classes), paste0("frac_", crop.classes$value[j], ".crops"))
    if (sum(idx) > 0) {
      # print(counter)
      str.idx[idx] <- counter
      counter <- counter + 1
    }
  }

  # Extract the crop values for placement purposes
  crop.val <- stringr::str_extract(rownames(extract.raster.classes), "\\d+")

  # Calculate the mean for ecah crop over all of the years
  raster.class.mean <- data.table("classes" = extract.raster.classes[, 1], "index" = str.idx, "crop.val" = crop.val) %>%
    group_by(index) %>%
    summarise(class.mean = mean(classes), crop.val = paste0("frac_", first(crop.val)))

  # get indices - for columns of outout - where classes match the output - create column names for datatable reference
  idx.col <- paste0("huc12.", class.possible.names[class.possible.names %in% raster.class.mean$crop.val])

  # Assign the extracted class fractions to the appropriate location in the output

  HUC.rast.class[idx.row, eval(idx.col) := as.list(raster.class.mean$class.mean)]
}

# Correct some column names that will be problematic later

# Classes column names updates - drop the HUC first column
HUC.rast.class <- HUC.rast.class %>%
  rename_with(function(x) str_replace_all(x, "[.]", "_")) %>%
  select(-huc12)

# Rename the number with the class name
colnames(HUC.rast.class)[colnames(HUC.rast.class) %in% eval(colnames(HUC.rast.class))] <- eval(convert.name)

idx.drop <- which(str_detect(colnames(HUC.rast.class), "^frac_$"))
HUC.rast.class <- HUC.rast.class %>% select(-eval(idx.drop))
```


```{r Save the data to a geopackage (OSG open source format) ,echo=TRUE}
sf::st_write(HUC.rast.class, "/Volumes/SHAG/PrestoGP_Covariates/AZO_HUC12_NASS_Crops.gpkg")
```
