---
title: "Calc_OLM_HUC10_Raster_Covariates"
author: "Kyle P Messier"
date: "5/19/23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required Libraries and Source Local Functions

```{r libraries and sourcing,echo=FALSE}

library(sf)
library(terra)
library(exactextractr)
library(lubridate)
library(tidyverse)
library(data.table)
library(nhdplusTools)


```

## Prepare Pesticide Data - sf format

```{r pesticide and NHD WBD data, echo=FALSE}
# Read in main pesticide data here
data.AZO <- sf::st_read("input/data_process/data_AZO_watershed_huc_join.shp")

# For efficient extraction, we just need the geometry
AZO.geometry <- sf::st_geometry(data.AZO)

# US bounding box
US.bb <- terra::ext(c(-124.7844079 , -66.9513812, 24.7433195, 49.3457868))

# NHD WBD Layer names
WBD.layers <- st_layers("input/WBD-National/WBD_National_GDB.gdb")

WBD <- st_read("input/WBD-National/WBD_National_GDB.gdb",layer = "WBDHU10") %>% 
  st_transform("EPSG:4326")

```


## Read in the OpenLandMaps (OLM) soil raster data

Here, we have an if-else statement. If the cropped (to US bounding box) have been
created, then we read those. Else, we read in the raw data and then crop. The process
of cropping takes a bit of time - about an hour - so we want to only do that once
if we can help it. 

We use the terra package to read and process rasters, which is the most efficient
method currently available in R for dealing with rasters.

```{r OLM raster data,echo=FALSE}
# Cropped Raster directory 
# cropped.raster.dir <- "/Volumes/SHAG/OpenLandMapData/OLM_Combined/"
cropped.raster.dir <- "input/OpenLandMapData/OLM_Combined/"

if ( length(list.files(path = cropped.raster.dir,
           pattern = "*.tif$", full.names = T))==7){ # If the cropped rasters have been generated,read those in
  
  pH.stack <- terra::rast(paste0(cropped.raster.dir,"OLM_US_Crop_pH.tif"))
  Clay_Content.stack <- terra::rast(paste0(cropped.raster.dir,"OLM_US_Crop_Clay_Content.tif"))
  Bulk_Density.stack <- terra::rast(paste0(cropped.raster.dir,"OLM_US_Crop_Bulk_Density.tif"))
  Sand_Content.stack <- terra::rast(paste0(cropped.raster.dir,"OLM_US_Crop_Sand_Content.tif"))
  Organic_Carbon.stack <- terra::rast(paste0(cropped.raster.dir,"OLM_US_Crop_Organic_Carbon.tif"))
  soil_order.stack <- terra::rast(paste0(cropped.raster.dir,"OLM_US_Crop_Soil_Order.tif"))
  texture.stack <- terra::rast(paste0(cropped.raster.dir,"OLM_US_Crop_Soil_Texture_Class.tif"))


}else{# otherwise generate them which takes a bit of time (~ 1 hour)
  
  # Directories of individual raw OLM data
OLM.dir.pH <- "input/OpenLandMapData/pH"
OLM.dir.Clay_Content <- "input/OpenLandMapData/Clay_Content"
OLM.dir.Bulk_Density <- "input/OpenLandMapData/Bulk_Density"
OLM.dir.Sand_Content <- "input/OpenLandMapData/Sand_Content"
OLM.dir.Organic_Carbon <- "input/OpenLandMapData/Organic_Carbon"
OLM.dir.Soil_Order <- "input/OpenLandMapData/Soil_Order"
OLM.dir.Texture <- "input/OpenLandMapData/USDA_Texture_Class"


# Get Raster data filenames/paths
OLM.filenames.pH <- list.files(path = OLM.dir.pH,
                            pattern = "*.tif", full.names = T)

pH.stack <- terra::rast(OLM.filenames.pH) %>% terra::crop(US.bb)
beepr::beep(sound =2)

#
OLM.filenames.Clay_Content <- list.files(path = OLM.dir.Clay_Content,
                               pattern = "*.tif", full.names = T)

Clay_Content.stack <- terra::rast(OLM.filenames.Clay_Content) %>% terra::crop(US.bb)
beepr::beep(sound =2)

#
OLM.filenames.Bulk_Density <- list.files(path = OLM.dir.Bulk_Density,
                                         pattern = "*.tif", full.names = T)

Bulk_Density.stack <- terra::rast(OLM.filenames.Bulk_Density) %>% terra::crop(US.bb)
beepr::beep(sound =2)

#
OLM.filenames.Sand_Content <- list.files(path = OLM.dir.Sand_Content,
                                         pattern = "*.tif", full.names = T)

Sand_Content.stack <- terra::rast(OLM.filenames.Sand_Content) %>% terra::crop(US.bb)
beepr::beep(sound =2)

# 
OLM.filenames.Organic_Carbon <- list.files(path = OLM.dir.Organic_Carbon,
                                         pattern = "*.tif", full.names = T)

Organic_Carbon.stack <- terra::rast(OLM.filenames.Organic_Carbon) %>% terra::crop(US.bb)
beepr::beep(sound =2)

# 
OLM.filenames.Soil_Order <- list.files(path = OLM.dir.Soil_Order,
                                         pattern = "*.tif", full.names = T)

soil_order.stack <- terra::rast(OLM.filenames.Soil_Order) %>% terra::crop(US.bb)

# 
OLM.filenames.Texture <- list.files(path = OLM.dir.Texture,
                                         pattern = "*.tif", full.names = T)

texture.stack <- terra::rast(OLM.filenames.Texture) %>% terra::crop(US.bb)


# Write the cropped rasters for future use and time saving
writeRaster(pH.stack,"input/OpenLandMapData/OLM_Combined/OLM_US_Crop_pH.tif")
writeRaster(Clay_Content.stack,"input/OpenLandMapData/OLM_Combined/OLM_US_Crop_Clay_Content.tif")
writeRaster(Bulk_Density.stack,"input/OpenLandMapData/OLM_Combined/OLM_US_Crop_Bulk_Density.tif")
writeRaster(Sand_Content.stack,"input/OpenLandMapData/OLM_Combined/OLM_US_Crop_Sand_Content.tif")
writeRaster(Organic_Carbon.stack,"input/OpenLandMapData/OLM_Combined/OLM_US_Crop_Organic_Carbon.tif")
writeRaster(soil_order.stack,"input/OpenLandMapData/OLM_Combined/OLM_US_Crop_Soil_Order.tif")
writeRaster(texture.stack,"input/OpenLandMapData/OLM_Combined/OLM_US_Crop_Soil_Texture_Class.tif")

}




# Combine the rasters in which we are calculating a mean value into one raster stack with
# many layers
OLM.stack.values <- c(pH.stack,Clay_Content.stack,Bulk_Density.stack,Sand_Content.stack,
                      Organic_Carbon.stack,soil_order.stack)

OLM.stack.classes <- texture.stack


# The soil texturer needs updating of its class names

texture_classes <- data.frame("classes" = c("clay","silty_clay","sandy_clay",
                     "clay_loam","silty_clay_loam","sandy_clay_loam", "loam",
                     "silt_loam","sandy_loam","silt",
                     "loamy_sand","sand"),"value" = 1:12)

for (i in 1:6){
  levels(OLM.stack.classes[[i]]) <- list(data.frame(ID = 1:12,
                                            textures = texture_classes$classes))
}

names(OLM.stack.classes) <- c("Texture_0cm","Texture_10cm","Texture_30cm",
                      "Texture_60cm","Texture_100cm","Texture_200cm")


```

Here, we use the terra extract function to extract the exact raster value
at each of our pesticide locations. First, we have to convert the pesticide 
data to a SpatVect (terra package), a requirement of the extract function. 

```{r Calculate exact grid points ,echo=TRUE}

# HUC10
HUC.rast.vals <- data.table("huc10" = data.AZO$huc10)

huc10.unique <- unique(data.AZO$huc10)

HUC.rast.vals[,paste0("huc10",".",names(OLM.stack.values)) := NA_real_]

HUC.rast.class <- data.table("huc10" = data.AZO$huc10)

class.df <- expand.grid(levels(OLM.stack.classes)[[1]]$ID, c("Texture_0cm","Texture_10cm","Texture_30cm",
                                                   "Texture_60cm","Texture_100cm","Texture_200cm"))

class.df <- class.df[order(class.df$Var1),]

class.possible.names <- paste0("frac_",class.df$Var1,".",class.df$Var2)


HUC.rast.class[,paste0("huc10",".",class.possible.names) := 0]

for (i in 1:length(huc10.unique)){
  print(i)
  
  # get index - for rows of output - where data match the given HUC10
  idx.row <- data.AZO$huc10 == huc10.unique[i]
  
  # get the given HUC10 geometry from the WBD data
  huc10.polygon <- dplyr::filter(WBD,huc10 == huc10.unique[i])
 
  # calculate the mean raster values in the HUC
  huc10.val <- exact_extract(OLM.stack.values,st_geometry(huc10.polygon),fun = "mean", stack_apply = TRUE)

  # Assign the extracted values to the appropriate location in the output
  HUC.rast.vals[idx.row,2:ncol(HUC.rast.vals)] <- huc10.val
  
  # calculate the fraction of each raster class in the HUC
  extract.raster.classes <- exact_extract(OLM.stack.classes,huc10.polygon,fun = "frac", stack_apply = TRUE)
  
  # get indexs - for columns of outout - where classes match the output
  idx.col <- which(class.possible.names %in% colnames(extract.raster.classes))+1

  # Assign the extracted class fractions to the appropriate location in the output
  HUC.rast.class[idx.row,idx.col] <- extract.raster.classes

}

#Correct some column names that will be problematic later

HUC.rast.vals <- HUC.rast.vals  %>% 
rename_with(function(x) str_replace_all(x,"[.]","_")) %>% 
rename_with(function(x) str_replace_all(x,"_{2}","_")) %>% 
rename_with(function(x) str_replace_all(x,"sol_order_usda_soiltax_","")) %>% 
rename_with(function(x) str_replace_all(x,"_1950_2017_v0_1","")) %>%
  select(-huc10)

# Classes column names updates
HUC.rast.class <- HUC.rast.class  %>% 
rename_with(function(x) str_replace_all(x,"[.]","_")) %>% 
  select(-huc10)

# Rename the number with the class name - ugly but it works
HUC.rast.class <- dplyr::rename_with(HUC.rast.class,~ gsub("frac_1_",paste0("frac_",texture_classes$classes[1],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_2_",paste0("frac_",texture_classes$classes[2],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_3_",paste0("frac_",texture_classes$classes[3],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_4_",paste0("frac_",texture_classes$classes[4],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_5_",paste0("frac_",texture_classes$classes[5],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_6_",paste0("frac_",texture_classes$classes[6],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_7_",paste0("frac_",texture_classes$classes[7],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_8_",paste0("frac_",texture_classes$classes[8],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_9_",paste0("frac_",texture_classes$classes[9],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_10_",paste0("frac_",texture_classes$classes[10],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_11_",paste0("frac_",texture_classes$classes[11],"_"),.x,fixed = TRUE)) %>%
  dplyr::rename_with(~ gsub("frac_12_",paste0("frac_",texture_classes$classes[12],"_"),.x,fixed = TRUE))

```


```{r Save the data to a geopackage (OSG open source format) ,echo=TRUE}

data.AZO.HUC10.OLM <- cbind(data.AZO,HUC.rast.vals,HUC.rast.class)
sf::st_write(data.AZO.HUC10.OLM,"input/Covariates/AZO_HUC10_OLM.gpkg")


```
